Gravitational Lens Recovery with \Glass: Measuring the mass profile and shape of a lens

### 2014-01-28 limits/ms.tex
Justin Changes

@@

\begin{abstract}
We use a new non-parametric gravitational modelling tool---\Glass---to
determine what quality of data (strong lensing, stellar kinematics, and/or
stellar masses) are required to measure the circularly averaged mass profile of
a lens and its shape. \Glass\ uses an under-constrained adaptive grid of mass
pixels to model the lens, searching through thousands of models to marginalise
over model uncertainties. Our key findings are as follows: (i) for pure lens
data, multiple sources with wide redshift separation give the strongest
constraints as this breaks the well-known mass-sheet or steepness degeneracy;
(ii) a single quad with time delays also performs well, giving a good recovery
of both the mass profile and its shape; (iii) stellar masses -- for lenses
where the stars dominate the central potential -- can also break the steepness
degeneracy, giving a recovery almost as good as having time delay data or
multiple source redshifts; (iv) stellar kinematics provide a robust measure of the mass at the half light radius of the stars $r \gg r_{1/2}$ that can also break the steepness degeneracy if $r_{1/2} \neq r_E$ -- the Einstein radius; and (v) if $r_E \sim r_{1/2}$, then stellar kinematic data can be used to probe the stellar velocity anisotropy 
$\beta(r)$ -- an interesting quantity in its own right. Where
information on the mass distribution from lensing and/or other probes becomes
redundant, this opens up the possibility of using strong lensing to constrain
cosmological models. We will study this, and present the first results from
\Glass\ applied to real data, in forthcoming papers.
\end{abstract}

@@

To see why lens modelling details are of crucial importance, let us recall the
essential quantities that appear in lensing.  First we have the
distances. Let $D_L$, $D_S$, $D_{LS}$ be the angular-diameter
distances to the lens, source, and from lens to source; these are all proportional to $c/H_0$ (where $c$ is the speed of light in vacuo, and $H_0$ the Hubble constant), with additional terms that depend on the particular choice of cosmology \citep[e.g.][]{1999coph.book.....P}. Typically:
\begin{equation}
D_L \approx z \frac c{H_0}\ \hbox{and}\ \frac{D_S}{D_{LS}} \sim 1.
\end{equation}
where the second relation follows from the fact that the light deflection angle is maximised when the lens is equidistant between the observer and the source \citep[e.g.][]{1992grle.book.....S}.

\hilight{GOT TO HERE. }

For multiple images the sky-projected density must exceed the critical
density of $D_S/D_{LS}$ times
\begin{equation}
\Sigma_L = \frac{c^2}{4\pi G D_L} \sim \frac{1\rm\,kg\,m^{-2}}{z_L}
\end{equation}
in some region.  The separation between the lensed images is of order
the Einstein radius $\theta_E$.  The image separation is related to
the mass by
\begin{equation}
\theta_E \sim \frac{R_G}{D_L} \frac{D_{LS}}{D_S}
\end{equation}
where $R_G$ is the gravitational radius $GM/c^2$.  If the source is a
quasar or otherwise rapidly variable, a time delay $\Delta t$ in the variability
will be present where
\begin{equation}
\Delta t \sim R_G/c \,.
\end{equation}
So in principle, one can not only measure the mass of the lens, one
can use the dependence on the cosmology-dependent $D$ factors to
extract the cosmological model and all its parameters.
\cite{1937ApJ....86..217Z} drew attention to the former, and
\cite{1964MNRAS.128..307R,1966MNRAS.132..101R} pointed out the latter,
all long before lenses were discovered.  The difficulty became
apparent soon after the discovery of the first lens by
\cite{1979Natur.279..381W}.  In the first ever paper on lens
modelling, \cite{1981ApJ...244..736Y} found that many plausible mass
distributions could reproduce the data.

\cite{1981ApJ...244..736Y} were remarkably prescient about the
subsequent development of lens modelling.  First, they introduced the
technique of choosing a parametric form for the lensing mass and then
fitting for the parameters, which is still the most common strategy
\citep[see for example][]{2010GReGr..42.2151K,2011A&ARv..19...47K}.
Second, they pointed out the non-uniqueness of lens models, or lensing
degeneracies.  Third, they suggested combining lensing data with
stellar kinematics and X-rays, to reduce the effect of the
degeneracies.  Later work, as well as following up these suggestions,
has introduced some ideas.  Four of these are important in the present
work.
\begin{itemize}
\item In free-form models, there is no specified parametric form for
  the mass distribution.  There are still assumptions (or priors) on
  the mass distribution, such as smoothness or being centrally
  concentrated
  \citep{1997MNRAS.292..148S,2005MNRAS.360..477D,2010ApJ...723.1678C}
  but these are much less restrictive than parametric forms.
  \cite{2006MNRAS.367.1209L} implemented a particularly elegant prior,
  which is that the mass distribution must be non-negative and no
  extra images are produced.  Free-form models are more commonly used with
  cluster lenses \citep{2009ApJ...690..154S,2013arXiv1304.2393S}, but
  can be used with galaxy lenses as well, where their less restrictive
  assumptions can be important.  For example, in time-delay galaxy
  lenses parametric models led to tensions with the Hubble parameter
  $H_0$ \citep[e.g.][]{2002astro.ph..4043K,2002ApJ...578...25K}, but
  the less restrictive assumptions of free-form models resolved those
  tensions \citep{2007ApJ...667..645R}.
\item Model ensembles, exploring a diverse range of possible mass
  distributions that nonetheless all fit the data, are a way of
  combating the non-uniqueness of models.  Model ensembles are
  possible in parametric models \citep{1999AJ....118...14B} but are
  more common in free-form models
  \citep{2000AJ....119..439W,2009ApJ...690..154S,2012MNRAS.425.3077L}.
\item Stellar kinematics has played a role in many modelling
  techniques.  The underlying reason is the virial theorem, which
  relates the line of sight velocity dispersion to the Einstein radius
  as
%
\begin{equation}
  \frac{\langle v^2_{\rm los}\rangle}{c^2} \approx
  \frac{\theta_E}{6\pi} \frac{D_S}{D_{LS}}
\end{equation}
%
  with the relation becoming exact for isothermal lenses.  The use of
  two-dimensional kinematics by \cite{2011MNRAS.415.2215B} is
  especially interesting.
\item The stellar mass in a lens can be inferred from photometry and
  compared with the total mass
  \citep{2005ApJ...623L...5F,2008MNRAS.383..857F,2011ApJ...740...97L}.
  Since the inferred stellar mass does depend on the assumed IMF,
  lenses in which stellar mass dominates can be used to derive upper
  bounds on the stellar $M/L$ \citep{2010MNRAS.409L..30F}. Lower bounds
  on stellar $M/L$ have also been claimed \citep{2013MNRAS.428.3183D}
  but these require upper-bound assumptions about the dark-matter
  profile.
\item Testing modelling strategies to see how well they recover
  simulated lenses is increasingly recognized as essential.  Simple
  blind tests have appeared in earlier work \citep[for example, Figure
    2 in][]{2000AJ....119..439W}, but more recently, tests against
  dynamically simulated galaxies are favoured
  \citep{2007ApJ...667..645R,2009MNRAS.393.1114B}.
\end{itemize}
There are also more modelling ideas in the literature.  One is to use
X-ray intensity and temperature profiles as a mass constraint
\citep[such as in][]{2013ApJ...765...25N}.  Another idea is to model
multiple lenses simultaneously, with one or more cosmological
parameters variable but shared between the lenses.  This strategy has
been used to constrain $H_0$ from time delays
\citep{2006ApJ...652L...5S,2008ApJ...679...17C,2010ApJ...712.1378P}.
In principle, cosmological parameters could be constrained even from a
single lens.  Some models have attempted this
\citep{2010Sci...329..924J,2013arXiv1306.4732S} starting from a prior
based on the values and uncertainties from the CMB; tests of this
technique simulated lenses are still awaited, in order to see how far
the results depend on the model assumptions.

In this paper, we use a new program---\Glass, the Gravitational Lensing AnalysiS Software---to determine what
combination of lensing, stellar mass, and stellar kinematic
constraints best constrain the projected mass profile and shape of several
gravitational lens produced from mock data.  The compute-intensive core of \Glass\ is a
parallel implementation of the high-dimensional sampling algorithm
from \cite{2012MNRAS.425.3077L}.  Around that core, \Glass\ has
extensible code (in software terminology, a framework) in Python,
encoding the lens equation, a basis and priors for the mass
distribution, and various post-processing and visualisation modules.  \Glass\
is described in \secref{sec:glass}.  In \secref{sec:theory} before that we
review the elements of lensing theory, stellar population synthesis, and
stellar dynamics we will need.

In \secref{sec:mockdata}, we describe our mock data.

In \secref{sec:results}, we present our results from applying \Glass\ to
these mock data.

\section{Formulation}\label{sec:theory}

@@

The lens equation
\begin{equation}
\begin{aligned}
\vec\beta &= \vec\theta - \frac{D_{LS}}{D_S}\vec\alpha(\vec\theta) \\
\vec\alpha(\vec\theta) &= \frac{4G}{c^2D_L} \int \Sigma(\vec\theta')
                          \frac{(\vec\theta - \vec\theta')}
                          {\ |\vec\theta - \vec\theta'|^2} \, d^2\vec\theta'
\end{aligned}
\label{eqn:lens_equation}
\end{equation}
maps an observed image position $\vec\theta$ to a source position
$\vec\beta$.  The lens can be thought of as a projected surface
density $\Sigma$ which diverts the path of a photon instantaneously
through the bending angle $\vec\alpha$.  The $D$ factors, as in the previous section, are angular
diameter distances, which depend on the particular choice of cosmological
parameters $\Omega$:
\begin{equation}
D_{LS} = \frac c{H_0} \frac{1+z_S}{1+z_L} \int_{z_L}^{z_S}
                      \frac{dz}{\sqrt{\Omega_m(1+z)^3 + \Omega_\Lambda}}
\end{equation}
%
and $D_L \equiv D_{0,L}$, $D_S \equiv D_{0,S}$.
One way to understand the lens equation is via Fermat's principle. We
can think of light as travelling only along extremum paths where
lensed images occur. Such paths occur at
the extrema of the photon {\it arrival time\/} $t(\vec\theta)$
that depends on the geometric path the photon takes and the general
relativistic gravitational time dilation due to a thin lens at
redshift $z_L$.
\begin{equation}
\begin{aligned}
\frac{ct(\vec\theta)}{(1+z_L)D_{L}}
&= {\textstyle\frac12} |\vec\theta - \vec\beta|^2
   \times \frac{D_{S}}{D_{LS}} \\
&- \frac{4GD_L}{c^2}
   \int \Sigma(\vec\theta') \ln |\vec\theta-\vec\theta'| \, d^2\vec\theta'
\label{full arrival time}
\end{aligned}
\end{equation}
The time here has been scaled to make it dimensionless.  We can make
the density dimensionless as well, by introducing
\begin{equation}
\kappa(\vec\theta) \equiv \frac{\Sigma(\vec\theta)}{\Sigma_L}
\end{equation}
and hence rewrite \eqnref{full arrival time} as
\begin{equation}
\tau(\vec\theta) = {\textstyle\frac12} |\vec\theta - \vec\beta|^2
                   \times \frac{D_{S}}{D_{LS}}
                 - \frac1\pi \int \kappa (\vec\theta')
                   \ln|\vec\theta - \vec\theta'| d^2\vec\theta'
\label{arrival time}
\end{equation}
The scaled arrival time $\tau$ is like a solid angle.  It is of order
the area (in steradians) of the full lensing system --- the expression
$\sim|\vec\theta - \vec\beta|^2$ is of order the image-separation
squared, and the other terms are of similar size.  For this reason, is
convenient to measure $\tau$ in arcsec$^{2}$.

Lensing observations provide information only at $\vec\theta$ where
there are images.  Hence, the arrival-time surface $\tau(\vec\theta)$
is not itself observable.  Its usefulness lies in that observables can
be derived from it.  Thus an image observed at $\vec\theta_1$ implies
that $\nabla\tau(\vec\theta_1)=0$.  A measurement of time delays
between images at $\theta_1$ and $\theta_2$ implies that
$t(\vec\theta_1)-t(\vec\theta_2)$ is known.  Interestingly, both these
types of observations give contraints that are linear in $\kappa$ and
$\vec\beta$.

The rather complicated dependence of lensing observables on the mass
distribution $\kappa(\vec\theta)$ has an important consequence: very different
mass distributions can result in similar observables.  This is the phenomenon
of lensing degeneracies.  While the non-uniqueness of lens models noted by
\cite{1981ApJ...244..736Y} already hinted at degeneracies, their existence was
first derived by \cite{1985ApJ...289L...1F}.  The most important is the
so-called mass-sheet degeneracy, which is that image positions remain invariant
if the $D_S/D_{LS} - \kappa(\vec\theta)$ \hilight{ $\leftarrow$ What is meant
by this? Can we just write $\tau(\vec\theta)$? }are each multiplied by an
arbitrary constant.  This amounts to multiplying the arrival-time surface by a
constant factor.  In fact there are infinitely many degeneracies
\citep{2000AJ....120.1654S} because any transformation of the arrival-time
surface away from the images has no effect on the lensing observables.  In
particular, there are degeneracies that involve the shape of the mass
distribution \citep{2006ApJ...653..936S,2013arXiv1306.4675S}.  Degeneracies
tend to be suppressed if there are sources at very different redshifts or
`redshift contrast' \citep{1998AJ....116.1541A,2009ApJ...690..154S}, because
the presence of different factors of $D_S/D_{LS}$ in the image plane makes it
more difficult to change the mass distribution and the arrival-time surface
without affecting the lensing observables.  But degeneracies are still present
with multiple source redshifts.  \citep{2008MNRAS.386..307L}.

@@

\begin{equation}
\sigma_p^2(R) = \frac{2}{I(R)}\int_R^\infty dr \left(1-\beta \frac{R^2}{r^2}\right) \frac{\nu \sigma_r^2 r}{\sqrt{r^2 - R^2}};
\label{eqn:sphericaljeans}
\end{equation}
\begin{equation} 
\sigma_r^2(r) = \frac{r^{-2\beta}}{\nu}\int_r^\infty r'^{2\beta} \nu \frac{GM(r')}{r'^2}dr'
\end{equation} 
where $\sigma_p$ is the projected velocity dispersion of the stars as a
function of projected radius $R$; $I(R)$ is the surface density of the stars;
$\nu(r)$ is the three dimensional stellar density; $\sigma_{r,t}(r)$ are the
radial and tangential velocity dispersions, respectively; $\beta(r) = 1 -
\sigma_t^2/2\sigma_r^2 = \mathrm{const.}$ is the velocity anisotropy (here
assumed to be constant, and not to be confused with $\vec\beta(\vec\theta)$ from
lensing); $G$ is Newton's gravitational constant; and $M(r)$ is
the mass profile that we would like to measure. By convension, we always write $R$ 
for a projected radius, and $r$ for a 3d radius.

It is immediately clear from \eqnref{eqn:sphericaljeans} that, even assuming
spherical symmetry, we have a degeneracy between the enclosed mass profile
$M(r)$ and the velocity anisotropy $\beta(r)$. This can be understood
intuitively since $\beta(r)$ measures the relative importance of radial versus
circular orbits and is intrinsically difficult to constrain given only one
component of the velocity vector for each star. Nonetheless, $\beta(r)$ can be
constrained given sufficiently many stars, since radial Doppler velocities
sample eccentric orbits as $r\rightarrow 0$ and tangential orbits as
$r\rightarrow \infty$ \citep[e.g.][]{2002MNRAS.330..778W}. It can also be
estimated if an independent measure of $M(r)$ is available -- for example
coming from strong lensing. Finally, we may integrate out the effect of unknown
$\beta(r)$ by performing a luminosity weighted average over all radii:

\begin{equation}
\overline{\sigma} = \frac{\int_0^{a_p} I(R) \sigma_p^2(R) 2\pi R dR}{\int_0^{a_p} I(R) 2\pi R dR}
\label{eqn:virialaverage}
\end{equation} 
where $a_p$ is the `aperture' over which the integral is performed. Ideally we
should take $a_p \rightarrow \infty$, however this is not practical for real
data where the surface brightness of the stars falls off rapidly with $a_p$. In
practice, $a_p > 2r_{1/2}$, where $r_{1/2}$ is the half light radius of the
stars, appears to be sufficient (see \S\ref{sec:results}). The above average
produces a single robust measure of the mass $\overline{\sigma}$ that is
independent of $\beta(r)$, provided that the stellar system is in a dynamic
quasi-equilibrium that obeys the virial theorem
\citep{2009ApJ...704.1274W,2010MNRAS.406.1220W,2012ApJ...754L..39A}. (Such an
averaging procedure was first used in a combined lensing/stellar kinematic
analysis in \citealt{2002MNRAS.337L...6T}.)

@@

\Glass is the Gravitational Lensing AnalysiS Software. It is an evolution of the concepts employed by the free form modelling tool
\PixeLens\ but completely rewritten in Python and C. 
The most compute intensive
portion was written in C but Python was chosen because of its flexibility as a 
language and for its large scientific library support. The flexibility allows 
\Glass\ to have quite sophisticated behavior while at the same time simplifying the 
user experience and reducing the overall development time. One of the striking
features is that the input file to \Glass\ is itself a Python program.
Understanding Python is not necessary for the most basic use, but this allows
a user to build complex analysis of a model directly into the input file. \Glass\
may also be used as an external library to other Python programs.  The software
is freely available for download or from the author%
%
\footnote{http://www.jpcoles.com.}.

>>>>

%For deprojection algorithm: 
\def\cth{\cos\theta}
\def\sth{\sin\theta}
\def\dth{{\rm d}\theta}

@@

\begin{abstract}
We use a new non-parametric gravitational modelling tool --- \Glass\ --- to
determine what quality of data (strong lensing, stellar kinematics, and/or
stellar masses) are required to measure the circularly averaged mass profile of
a lens and its shape. \Glass\ uses an under-constrained adaptive grid of mass
pixels to model the lens, searching through thousands of models to marginalise
over model uncertainties. Our key findings are as follows: (i) for pure lens
data, multiple sources with wide redshift separation give the strongest
constraints as this breaks the well-known mass-sheet or steepness degeneracy;
(ii) a single quad with time delays also performs well, giving a good recovery
of both the mass profile and its shape; (iii) stellar masses -- for lenses
where the stars dominate the central potential -- can also break the steepness
degeneracy, giving a recovery almost as good as having time delay data or
multiple source redshifts; (iv) stellar kinematics provide a robust measure of the mass at the half light radius of the stars $r_{1/2}$ that can also break the steepness degeneracy if $r_{1/2} \neq r_E$ -- the Einstein radius; and (v) if $r_E \sim r_{1/2}$, then stellar kinematic data can be used to probe the stellar velocity anisotropy 
$\beta$ -- an interesting quantity in its own right. Where
information on the mass distribution from lensing and/or other probes becomes
redundant, this opens up the possibility of using strong lensing to constrain
cosmological models. We will study this, and present the first results from
\Glass\ applied to real data, in forthcoming papers.
\end{abstract}

@@

To see why lens modelling details are of crucial importance, let us recall the
essential quantities that appear in lensing.  First we have the
distances. Let $D_L$, $D_S$, $D_{LS}$ be the angular-diameter
distances to the lens, source, and from lens to source; these are all proportional to $c/H_0$ (where $c$ is the speed of light in vacuo, and $H_0$ the Hubble constant), with additional terms that depend on the particular choice of cosmology \citep[e.g.][]{1999coph.book.....P}. Typically:
\begin{equation}
D_L \approx z \frac c{H_0}\ \hbox{and}\ \frac{D_S}{D_{LS}} \sim 1.
\end{equation}

For multiple images, the sky-projected density must exceed the critical
density \citep[e.g.][]{1992grle.book.....S}:

\begin{equation}
\Sigma_{\rm crit} = \frac{D_S}{D_{LS}} \Sigma_L = \frac{D_S}{D_{LS}} \frac{c^2}{4\pi G D_L} \sim \frac{1\rm\,kg\,m^{-2}}{z_L}
\end{equation}
in some region.  The separation between the lensed images is of order
the Einstein radius $\theta_E$, which is related to
the mass by:
\begin{equation}
\theta_E \sim \frac{R_G}{D_L} \frac{D_{LS}}{D_S}
\end{equation}
where $R_G = GM/c^2$ (with $M$ the projected mass enclosed within $\theta_E$) is the gravitational radius. If the source is a
quasar or otherwise rapidly variable, a time delay $\Delta t$ in the variability
will be present where:
\begin{equation}
\Delta t \sim R_G/c \,.
\end{equation}
So in principle, one can not only measure the mass of the lens, one
can use the dependence on the cosmology-dependent $D$ factors to
extract the cosmological model and all its parameters.
\cite{1937ApJ....86..217Z} drew attention to the former, and
\cite{1964MNRAS.128..307R,1966MNRAS.132..101R} pointed out the latter,
all long before lenses were discovered.  The difficulty with actually doing this, however, became
apparent soon after the discovery of the first lens by
\cite{1979Natur.279..381W}.  In the first ever paper on lens
modelling, \cite{1981ApJ...244..736Y} found that many plausible mass
distributions could reproduce the data.

\cite{1981ApJ...244..736Y} were remarkably prescient about the
subsequent development of lens modelling.  First, they introduced the
technique of choosing a parametric form for the lensing mass and then
fitting for the parameters, which is still the most common strategy
\citep[see for example][]{2010GReGr..42.2151K,2011A&ARv..19...47K}.
Second, they pointed out the non-uniqueness of lens models -- lensing
degeneracies.  Third, they suggested combining lensing data with
stellar kinematics and X-rays, to reduce the effect of the
degeneracies.  Later work, as well as following up these suggestions,
has introduced some further new ideas.  Four of these are important for the present
work:
\begin{enumerate}
\item {\bf Free-form modelling:} In `free-form' or non-parametric modelling, there is no specified parametric form for
  the mass distribution.  There are still assumptions (or priors) on
  the mass distribution, such as smoothness or being centrally
  concentrated
  \citep{1997MNRAS.292..148S,2005MNRAS.360..477D,2009A&A...500..681M,2010ApJ...723.1678C}
  but these are much less restrictive than parametric forms (\cite{2006MNRAS.367.1209L} implemented a particularly elegant prior, which is that the mass distribution must be non-negative and no
  extra images are produced). To be concrete, we define from here on:

  \begin{quote}
  {\it Non-parametric, or `free-form' $\equiv$ more parameters than data constraints (i.e. deliberately under-constrained)}
  \end{quote}

  Being under-constrained, it is then {\it necessary} to explore model degeneracies rather than find a single `best-fit' solution. Free-form models are more commonly used with
  cluster lenses \citep{2006ApJ...652L...5S,2009ApJ...690..154S,2009A&A...500..681M,2013arXiv1304.2393S}, but
  can be used with galaxy lenses as well, where their less restrictive
  assumptions can be important.  For example, in time-delay galaxy
  lenses, parametric models have historically led to tension with independent measures of the Hubble parameter
  $H_0$ \citep[e.g.][]{2002astro.ph..4043K,2002ApJ...578...25K}; these are resolved once the less restrictive assumptions of free-form models are permitted \citep{2007ApJ...667..645R}.
\item Model ensembles, exploring a diverse range of possible mass
  distributions that nonetheless all fit the data, are a way of
  combating the non-uniqueness of models.  Such ensembles are
  possible in parametric models \citep{1999AJ....118...14B} but are
  more common in free-form models
  \citep{2000AJ....119..439W,2009ApJ...690..154S,2012MNRAS.425.3077L}.
\item {\bf Stellar kinematic constraints:} This was first suggested by \citet{2002MNRAS.337L...6T} as a means to break lensing degeneracies and determine cosmological parameters. This works because stellar kinematics can provide an independent estimate of the Einstein radius, via the virial theorem:
%
\begin{equation}
  \frac{\langle v^2_{\rm los}\rangle}{c^2} \approx
  \frac{\theta_E}{6\pi} \frac{D_S}{D_{LS}}
\label{eqn:virial}
\end{equation}
 where $\langle v^2_{\rm los}\rangle$ is the projected stellar velocity dispersion, and the above relation becomes exact for isothermal lenses.  The use of
  two-dimensional kinematics by \cite{2011MNRAS.415.2215B} is
  especially interesting.
\item {\bf Stellar mass constraints:} The stellar mass in a lens can be inferred from photometry and
  compared with the total mass
  \citep{2005ApJ...623L...5F,2008MNRAS.383..857F,2011ApJ...740...97L}.
  Since the inferred stellar mass depends on the assumed IMF,
  lenses in which stellar mass dominates can be used to derive upper
  bounds on the stellar $M/L$ \citep{2010MNRAS.409L..30F}. Lower bounds
  on stellar $M/L$ have also been claimed \citep{2013MNRAS.428.3183D}
  but these require upper-bound assumptions about the dark-matter
  profile. \hilight{JR. I feel like we need more citations here and throughout. We must be skipping over an Armada of literature.}
\item {\bf Testing modelling strategies:} Using mock data to see how well a given model can recover
  simulated lenses is increasingly being recognised as essential.  Simple
  blind tests have appeared in earlier work \citep[for example, Figure
    2 in][]{2000AJ....119..439W}, but more recently, tests against
  dynamically simulated galaxies or clusters are favoured
  \citep{2007ApJ...667..645R,2007MNRAS.380.1729L,2009A&A...500..681M,2009MNRAS.393.1114B,2010ApJ...723.1678C}.
\end{enumerate}

There are two further key modelling ideas in the literature that we will not touch upon in this present work: to use
X-ray intensity and temperature profiles as a mass constraint
\citep[e.g.][]{2013ApJ...765...25N}; and to model
multiple lenses simultaneously, with one or more cosmological
parameters variable but shared between the lenses. This latter strategy has
been used to constrain $H_0$ from time delay lenses 
\citep{2006ApJ...652L...5S,2008ApJ...679...17C,2010ApJ...712.1378P}.
(In principle, cosmological parameters can be constrained even from a
single lens if the data are good enough. Some recent models have attempted this starting from a prior
based on the Cosmic Microwave Background (CMB) radiation \citep{2010Sci...329..924J,2013arXiv1306.4732S}.)

In this paper, we use a new program --- \Glass\ (Gravitational Lensing AnalysiS Software) --- to determine what
combination of lensing, stellar mass, and stellar kinematic
constraints best constrain the projected mass profile and shape of several
gravitational lens produced from mock data.  The compute-intensive core of \Glass\ is a
parallel implementation of the high-dimensional sampling algorithm
from \cite{2012MNRAS.425.3077L}.  Around that core, \Glass\ has
extensible code (in software terminology, a framework) in Python,
encoding the lens equation, a basis and priors for the mass
distribution, and various post-processing and visualisation modules.  \Glass\
is described in \secref{sec:glass}.  In \secref{sec:theory}, we
review the key elements of lensing theory, stellar population synthesis, and
stellar dynamics we will need. In \secref{sec:mockdata}, we describe our mock data. In \secref{sec:results}, we present our results from applying \Glass\ to these mock data. Finally, in \S\ref{sec:conclusions} we present our conclusions.

\section{Theory}\label{sec:theory}

@@

The lens equation:
\begin{equation}
\begin{aligned}
\vec\beta &= \vec\theta - \frac{D_{LS}}{D_S}\vec\alpha(\vec\theta) \\
\vec\alpha(\vec\theta) &= \frac{4G}{c^2D_L} \int \Sigma(\vec\theta')
                          \frac{(\vec\theta - \vec\theta')}
                          {\ |\vec\theta - \vec\theta'|^2} \, d^2\vec\theta'
\end{aligned}
\label{eqn:lens_equation}
\end{equation}
maps an observed image position $\vec\theta$ to a source position
$\vec\beta$.  The lens can be thought of as a projected surface
density $\Sigma$ which diverts the path of a photon instantaneously
through the bending angle $\vec\alpha$.  The $D$ factors, as in the previous section, are angular
diameter distances, which depend on the particular choice of cosmological
parameters $\Omega$:
\begin{equation}
D_{LS} = \frac c{H_0} \frac{1+z_S}{1+z_L} \int_{z_L}^{z_S}
                      \frac{dz}{\sqrt{\Omega_m(1+z)^3 + \Omega_\Lambda}}
\end{equation}
%
and $D_L \equiv D_{0,L}$, $D_S \equiv D_{0,S}$.
One way to understand the lens equation is via Fermat's principle. We
can think of light as travelling only along extremum paths where
lensed images occur. Such paths occur at
the extrema of the photon {\it arrival time\/} $t(\vec\theta)$
that depends on the geometric path the photon takes and the general
relativistic gravitational time dilation due to a thin lens at
redshift $z_L$:
\begin{equation}
\begin{aligned}
\frac{ct(\vec\theta)}{(1+z_L)D_{L}}
&= {\textstyle\frac12} |\vec\theta - \vec\beta|^2
   \times \frac{D_{S}}{D_{LS}} \\
&- \frac{4GD_L}{c^2}
   \int \Sigma(\vec\theta') \ln |\vec\theta-\vec\theta'| \, d^2\vec\theta'
\label{full arrival time}
\end{aligned}
\end{equation}
We can simplify the above equation by introducing a dimensionless time $\tau$ and density $\kappa$:

\begin{equation}
\tau(\vec\theta) = \frac{ct(\vec\theta)}{(1+z_L)D_{L}} \,\,\,;\,\,\, \kappa(\vec\theta) \equiv \frac{\Sigma(\vec\theta)}{\Sigma_L}
\end{equation}
and hence rewrite \eqnref{full arrival time} as:
\begin{equation}
\tau(\vec\theta) = {\textstyle\frac12} |\vec\theta - \vec\beta|^2
                   \times \frac{D_{S}}{D_{LS}}
                 - \frac1\pi \int \kappa (\vec\theta')
                   \ln|\vec\theta - \vec\theta'| d^2\vec\theta'
\label{arrival time}
\end{equation}
The scaled arrival time $\tau$ is like a solid angle. It is of order
the area (in steradians) of the full lensing system. The expression
$|\vec\theta - \vec\beta|^2$ is of order the image-separation
squared, and the other terms are of similar size.  For this reason, is
convenient to measure $\tau$ in arcsec$^{2}$.

Lensing observations provide information only at $\vec\theta$ where
there are images.  Hence, the arrival-time surface $\tau(\vec\theta)$
is not itself observable.  Its usefulness lies in that observables can
be derived from it.  An image observed at $\vec\theta_1$ implies
that $\nabla\tau(\vec\theta_1)=0$.  A measurement of time delays
between images at $\theta_1$ and $\theta_2$ implies that
$t(\vec\theta_1)-t(\vec\theta_2)$ is known.  Interestingly, both these
types of observations give contraints that are linear in $\kappa$ and
$\vec\beta$.

The rather complicated dependence of lensing observables on the mass
distribution $\kappa(\vec\theta)$ has an important consequence: very different
mass distributions can result in similar observables.  This is the phenomenon
of lensing degeneracies.  While the non-uniqueness of lens models noted by
\cite{1981ApJ...244..736Y} already hinted at degeneracies, their existence was
first derived by \cite{1985ApJ...289L...1F}.  The most important is the
so-called mass-sheet degeneracy, which is that image positions remain invariant
if $\tau(\vec\theta)$ is multiplied by an
arbitrary constant.  This corresponds to rescaling the surface density at the images $\kappa(\vec\theta)$. In fact there are infinitely many degeneracies
\citep{2000AJ....120.1654S} because any transformation of the arrival-time
surface away from the images has no effect on the lensing observables.  In
particular, there are degeneracies that involve the shape of the mass
distribution \citep{2006ApJ...653..936S,2013arXiv1306.4675S}.  Degeneracies
tend to be suppressed if there are sources at very different redshifts or
`redshift contrast' \citep{1998AJ....116.1541A,2009ApJ...690..154S}, because
the presence of different factors of $D_S/D_{LS}$ in the image plane makes it
more difficult to change the mass distribution and the arrival-time surface
without affecting the lensing observables.  But degeneracies are still present
with multiple source redshifts.  \citep{2008MNRAS.386..307L}.

@@

\begin{equation}
\sigma_p^2(R) = \frac{2}{I(R)}\int_R^\infty dr \left(1-\beta \frac{R^2}{r^2}\right) \frac{\nu \sigma_r^2 r}{\sqrt{r^2 - R^2}};
\label{eqn:sphericaljeans}
\end{equation}
\begin{equation} 
\sigma_r^2(r) = \frac{r^{-2\beta}}{\nu}\int_r^\infty r'^{2\beta} \nu \frac{GM_{\rm 3D}(r')}{r'^2}dr'
\end{equation} 
where $\sigma_p$ is the projected velocity dispersion of the stars as a
function of projected radius $R$; $I(R)$ is the surface density of the stars;
$\nu(r)$ is the three dimensional stellar density; $\sigma_{r,t}(r)$ are the
radial and tangential velocity dispersions, respectively; $\beta(r) = 1 -
\sigma_t^2/2\sigma_r^2 = \mathrm{const.}$ is the velocity anisotropy (here
assumed to be constant, and not to be confused with $\vec\beta(\vec\theta)$ from
lensing); $G$ is Newton's gravitational constant; and $M_{\rm 3D}(r)$ is
the mass profile that we would like to measure. By convension, we always write $R$ 
for a projected radius, and $r$ for a 3D radius.

It is immediately clear from \eqnref{eqn:sphericaljeans} that, even assuming
spherical symmetry, we have a degeneracy between the enclosed mass profile
$M_{\rm 3D}(r)$ and the velocity anisotropy $\beta(r)$. This can be understood
intuitively since $\beta(r)$ measures the relative importance of radial versus
circular orbits and is intrinsically difficult to constrain given only one
component of the velocity vector for each star. Nonetheless, $\beta(r)$ can be
constrained given sufficiently many stars, since radial Doppler velocities
sample eccentric orbits as $r\rightarrow 0$ and tangential orbits as
$r\rightarrow \infty$ \citep[e.g.][]{2002MNRAS.330..778W}. It can also be
estimated if an independent measure of $M_{\rm 3D}(r)$ is available -- for example
coming from strong lensing.

While $M_{\rm 3D}(r)$ is difficult to measure from stellar kinematics alone, the mass within the half light radius is robustly recovered \citep[e.g.][]{2009ApJ...704.1274W,2010MNRAS.406.1220W,2012ApJ...754L..39A} since stellar systems in dynamic quasi-equilibrium obey the virial theorem (equation \ref{eqn:virial}). This means that stellar kinematics can break the steepness degeneracy if $r_{1/2} \neq r_E$, where $r_E = D_L \theta_E$ is the physical Einstein radius. We test this expectation in \S\ref{sec:results}.

@@

\Glass\ is the Gravitational Lensing AnalysiS Software. It is an evolution of the concepts employed by the free form modelling tool
\PixeLens\ but completely rewritten in Python and C. 
The most compute intensive
portion was written in C but Python was chosen because of its flexibility as a 
language and for its large scientific library support. The flexibility allows 
\Glass\ to have quite sophisticated behavior while at the same time simplifying the 
user experience and reducing the overall development time. One of the striking
features is that the input file to \Glass\ is itself a Python program.
Understanding Python is not necessary for the most basic use, but this allows
a user to build complex analysis of a model directly into the input file. \Glass\
may also be used as an external library to other Python programs.  The software
is freely available for download or from the author%
%
\footnote{http://www.jpcoles.com.}.
